Unsupervised Acquisition of Comprehensive Multiword Lexicons \\ using Competition in an $n$-gram Lattice

### 2017-01-19 tacl.tex
Edits to end of Section 2 + added \newcite

\begin{document}
\maketitle
\begin{abstract}

We present a new model for acquiring unrestricted multiword lexicons from large corpora based on  competition among $n$-gram candidates. In contrast to the standard approach of simple ranking by association measure, in our model $n$-grams are arranged in a lattice structure based on subsumption and overlap relationships, with nodes inhibiting other nodes in their vicinity when they are selected as a lexical item. We show how the configuration of such a lattice can be optimized tractably, and our evaluation using annotations of sampled $n$-grams indicates that it is consistently outperforms alternatives by at least 0.05 f-score across several corpora and languages.

\end{abstract}

@@

Despite over 25 years of research in computational linguistics aimed at acquiring multiword lexicons using corpora statistics, and growing evidence that speakers both produce and process language primarily in terms of memorized sequences \cite{Wray08}, the individual word nonetheless stubbornly remains the \textit{de facto} standard unit for most research in modern NLP. The potential of multiword knowledge to improve both the automatic processing of language as well as offer new understanding of human acquisition and usage of language is the primary motivator of this work. We present a effective, expandable, and above all tractable new approach in comprehensive multiword lexicon acquisition that aims to find a middle ground between standard MWE acquisition approaches based on association measures \cite{Ramisch14}, and more sophisticated statistical models \cite{Newman12} which fail to scale to the large corpora which are now the main sources of the distributional information which are central to lexical knowledge in modern NLP systems.

@@

In this paper we will refer to the targets of our lexicon creation efforts as \termdef{formulaic sequences}, following the terminology of Wray \shortcite{Wray02,Wray08}, wherein a formulaic sequence (FS) is defined as ``a sequence, continuous or discontinuous, of words or other elements, which is, or appears to be, prefabricated: that is, stored and retrieved whole from memory at the time of use, rather than being subject to generation or analysis by the language grammar.'' In other words, a formulaic sequence shows signs of being part of a mental lexicon. Though by this definition single individuals or small groups may have their own FS, in this paper, we are only interested in FS that are shared by a recognizable language community. In computational linguistics, the most common term used to describe multiword lexical units is \emph{multiword expression} or ``MWE'' \cite{Sag02,Baldwin10}, but here we wish to make a principled distinction between at least somewhat non-compositional, strongly lexicalized MWEs and FS, a near superset which includes most MWEs but also (apparently) compositional linguistic formulas. This distinction is not a new one; it exists, for example, in the original paper of \newcite{Sag02} in the distinction between lexicalized and institutionalized phrases, and also to some extent in the MWE annotation of \newcite{Schneider14a}, who distinguish between weak (collocational)\footnote{Here we avoid the term \emph{collocation} entirely due to even more confusion with respect to in its interpretation. Though in some contexts it has been used in way that is similar to our definition of FS, it can be applied to any words that show a statistical tendency to appear in the vicinity of one another for any reason: for instance, the pair of words \ex{doctor/nurse} might be considered a collocation \cite{Ramisch14}}  and strong (non-compositional) MWEs. It is our contention, however, that separate, precise terminology is useful for research targeted at either class: we need not strain the concept of MWE to include items which do not require special semantics, nor are we inclined to disregard the larger formulaticity of language simply because it is not the dominant focus of MWE research. Many MWE researchers would defensibly balk at including in their MWE lexicons and corpus annotations (English) formulaic sequences such as \ex{there is something going on}, \ex{it is more important than ever to \dotts}, \ex{\dotts do not know what it is like to \dotts}, \ex{there is no shortage of\dotts}, \ex{the rise and fall of\dotts}, \ex{now is not the time to\dotts}, \etc as well as tens of thousands of other such phrases which, along with less compositional MWEs like \ex{be worth \gap weight in gold}, fall under the FS umbrella. Another reason to consider a different terminology is that there are classes of what is typically considered MWE that don't fit well into an FS framework, for instance novel compound nouns whose semantics are accessible by analogy (\eg \ex{plastic house}). Also, we exclude from the definition of both FS and MWE those named entities which refer to people or places which are little-known and/or whose surface form appears derived (\eg \ex{Mrs. Barbara W. Smith} or \ex{Smith Garden Supplies Ltd}). \figref{fig:terminology} show the conception of the relationship between FS, MWE and (multiword) named entities that we assume for this paper.

@@

Regardless of the terminology used to describe them, the starting point for multiword lexicon creation has been typically lexical association measures \cite{Church90,Dunning93,Schone01,Evert04,Pecina10,DeAraujo11,Kulkarni11,Ramisch14}; when these methods are used to build a lexicon, particular syntactic patterns and thresholds for the metrics are typically chosen. Only some of these measures generalize tractably beyond two words, for example PMI \cite{Church90}, the log ratio of the joint probability to the product of the marginal probabilities of the individual words. Other measures specifically designed to address sequences of larger than two words include the $c$-value \cite{Frantzi00}, a metric designed for term extraction which weights term frequency by the log length of the $n$-gram while penalizing $n$-grams that appear in frequent larger ones, and mutual expectation \cite{Dias99}, which produces a normalized statistic that reflects how much a candidate phrase resists the omission of any particular word. Similarly, the lexical predictability ratio (LPR) of Brooke et \al \cite{Brooke15b} is a association measure intended for any possible syntactic pattern which is calculated by discounting syntactic predictability (as provided by part-of-speech information) from the overall conditional probability for each word given the other words in the phrase. Though most association measures involve only usage statistics of the phrase and its subparts, the DRUID measure is an exception which uses distributional semantics around the phrase to identify how easily $n$-gram could be replaced by a single word \cite{Riedl15}.

Typically multiword lexicons are created by ranking $n$-grams according to an association measure and applying a threshold. The algorithm of Silva and Lopes \shortcite{Silva99} is somewhat more sophisticated, in that it identifies the local maxima of association measures across subsuming $n$-grams within a sentence to identify MWEs of unrestricted length; its effectiveness beyond noun phrases, however, seems relatively limited \cite{Ramisch12}. Brooke et \al \shortcite{Brooke14a,Brooke15b} developed a heuristic method intended for general FS extraction in larger corpora, first using conditional probably statistics to do an initial (single pass) coarse-grained segmentation of the corpus, followed by a pass through the resulting vocabulary, breaking larger units into smaller ones based on tradeoff between marginal and conditional statistics (\ie LPR). Beyond association measures, other general unsupervised approaches to the multiword unit identification include that of Newman et \al \shortcite{Newman12}, who used a generative Dirichlet Process model which jointly creates a linear segmentation of the corpus and a multiword vocabulary of keyphrases. Gimpel and Smith \shortcite{Gimpel11} focus specifically on deriving word sequences with gaps using a generative model, with the intent of improving machine translation. A general drawback to these generative methods, relative to approaching involving association metrics, is scalability and lack of determinacy.

Other research in MWEs has tended to be rather focused on particular syntactic patterns, looking at, for instance, just verb/noun combinations \cite{Fazly09}. The work of Schneider et \al \shortcite{Schneider14b} is a rare example of a comprehensive token-level MWE identification system which distinguishes a full range of MWE sequences in the English Web Treebank, including those involving gaps, using a supervised sequence tagging model; like other models in this space, Schneider et \al make use of existing manual lexical resources and they note that an (unsupervised) automatic lexical resource could be useful addition to the model, though attempts to do so have had mixed results \cite{Riedl16}.

The motivation for building lexicons of formulaic sequences naturally overlaps of those for MWE: models of distributional semantics, in particular, can benefit from sensitivity to multiword units (citation needed?????????), as can parsing \cite{Constant16}. One major motivation for looking beyond MWEs is the ability to carry out broader linguistic analyses. Within corpus linguistics, multiword sequences have studied in the form of \textit{lexical bundles} \cite{Biber04}, which are simply $n$-grams that occur above a certain frequency threshold: in order to get very reliable phrases, the threshold is typically set rather high (Biber et \al use 40 occurrences in 1 million words). Like FS, Lexical bundles generally involve larger phrasal chunks that would be missed by traditional MWE extraction, and so research in this area has tended to focus on how particular formulaic phrases (\eg \textit{if you look at}) are indicative of particular genres (\eg university lectures). Lexical bundles have been applied, in particular, to learner language: for example, Chen and Baker \shortcite{Chen10} showed that non-native student writers use severely a restricted range of lexical bundle types, and tend to overuse those types, while Granger and Bestgen \shortcite{Granger14} investigate the role of proficiency, demonstrating that intermediate learners underuse lower-frequency bigrams and overuse high-frequency bigrams relative to advanced learners. Sakaguchi et \al \shortcite{Sakaguchi16} demonstrate that improving fluency (which is closely linked to the use of linguistic formulas) is more important than improving strict grammaticality with respect to native speaker judgments of non-native productions; Brooke et \al \shortcite{Brooke15b} explcitly argues for FS lexicons as a way to identify, track, and improve learner proficiency.

@@

The first step in the process is derive a set of $n$-grams and related statistics from a large, unlabeled corpus of text. Since our primary association measure is an adaption of LPR, our approach in this subsection mostly follows Brooke et \al \shortcite{Brooke15b} up until the last stage. An initial requirement of any such method is a $n$-gram frequency threshold, which for us (like Brooke et \al 2015) is 1 instance per 10 million words; we have done a manual analysis using the MWE corpus of Schneider et \al \shortcite{Schneider14a} which indicates very good (over 90\%) type-level MWE coverage using the frequency filtered $n$-gram statistics from the ICWSM blog corpus (see \secref{sec:evaluation}), excepting those highly particular proper names which we consider outside the scope of FS identification. Note that our set of $n$-grams also include gapped or non-contiguous $n$-grams; the definition of gaps is language specific---we will discuss the gaps for individual languages further in \secref{sec:evaluation}---but the general idea (and this is confirmed by empirical analysis) is that for any particular language there is a relatively small set of syntactic types of limited length that will tend to appear as a gap inside a formulaic sequence. In a part-of-speech tagged corpus, potential gaps can be identified by a part-of-speech regular expression. Putting in such a restriction is useful because it greatly lowers the number of gapped $n$-gram types, increasing precision and improving speed. Another way the number of types are lowered is by excluding punctuation and by lemmatizing.  Note that as long as the count threshold is significantly above 1, efficient extraction of all $n$-grams can be done iteratively ($n$ passes through the corpus) for both types in combination.

@@

In the segmentation approach of Brooke et \al \shortcite{Brooke15b}, LPR for an entire span is calculated as a product of the individual LPRs, but here we will use the minimum LPR across the words in the sequence (minLPR):

@@

\begin{displaymath}
E_0(t) = minLPR(t)^{-d_0(t)}
\end{displaymath}
\noident

And $E_1$ is the cost parameter $C$ taken to the negative power of an analogous function $d_1$:

\begin{displaymath}
E_1(t) = C^{-d_1(t)}
\end{displaymath}
\noident

The values of $d_0$ and $d_1$ are dependent on the context of the node in the lattice, but are both 1 when a node is not being influenced by other nodes. They are also restricted to always being positive.

@@

\begin{displaymath}
clear(t) = \prod_{i=1}^{m}{ min(1,\frac{minLPR(t)}{minLPR(u_i)})}
\end{displaymath}
where $u_i$ is the $i$th clearing node of $t$. We refer to this mechanism as clearing because it tends to clear away a variety of trivial uses of common FS which many have higher LPR due to the lexical and syntactic specificity of the FS. For instance, in \figref{fig:example} if the node \ex{keep \gap under wraps}  is turned on and has a minLPR of 8, then, if the minLPR of a node such as \ex{keep \gap under wraps for} is 4, clear(t) will be 0.5. We can now define d_{0}(t) as:

@@

We evaluate our approach across three different languages including evaluation sets derived from four different corpora. In English, we follow Brooke et \al \shortcite{Brooke15b} in using a 890M token filtered portion of the ICWSM social media corpus \cite{ICWSM} tagged with the Tree Tagger \cite{Schmid95}. To facilitate a comparison with Newman et \al \shortcite{Newman12}, which does not scale up to a corpus as large as the ICWSM, we also build a lexicon using the 100M token British National Corpus \cite{BNC}, using the standard CLAWS-derived POS tags for the corpus. Lemmatization included removing all inflectional marking from both words and POS tags. For English, we took the same POS regex used in Brooke et \al \shortcite{Brooke15b}, which includes simple nouns and portions thereof, up to a maximum length of 4.

@@

Brooke et \al \shortcite{Brooke15b} introduced a method for evaluating formulaic sequence extraction without a reference lexicon or direct annotation of the output of a model: instead, $n$-grams are sampled after applying the frequency threshold and annotated as being either an FS or not, allowing for calculation of a true F-score for any model. We use the annotation of 2000 $n$-grams in the ICWSM corpus from that earlier work, and applied the same annotation methodology to the other corpora discussed above: based on written guidelines derived from the definitions of Wray \cite{Wray08} and at least one practice round,three native-speaker, educated annotators judged 500 contiguous $n$-grams and another 500 gapped $n$-grams for each of the new corpora.

Other than the inclusion of new languages, our test sets differ from Brooke et \al \shortcite{Brooke15b} in two ways. First, instead of relying on strict majority annotation, we entirely excluded $n$-grams which just one annotator marked as FS. One advantage of a type-based annotation approach (as compared to a corpus annotation), particularly with regards to an task with a known subjective component, is that it is quite sensible to simply discard borderline cases, improving the reliability of evaluation at the cost of some representativeness.  Second, for the main evaluation we collapsed gapped and contiguous $n$-grams into a single test set. The rationale is that the number of positive gapped examples was too low (particularly for the BNC and the Croatian corpus) to provide a reliable independent F-score (see further of discussion of gapped phrases in \secref{sec:discussion}). Statistics for the four test sets are given in \tabref{tab:stats}.

@@

Our primary comparison is with the heuristic LPR model from Brooke et \al \shortcite{Brooke15b}, which is scalable to large corpora and includes gapped $n$-grams. For the BNC, we also compare, separately, the DP-seg model from Newman et \al \shortcite{Newman12} with recommended settings, and the LocalMaxs algorithm of Silva and Lopes \shortcite{Silva99} using the SCP as the association measure; because these other approaches only generate sequential multiword units, we use only the sequential part of the BNC test set. All comparison approaches have themselves been compared against a wide range of association measures in earlier work;\footnote{One popular association measure which is theoretically extensible to $n$-grams of any length but which has not been addressed in the context of unrestricted MWE/FS identification is the (log-)likelihood ratio \cite{Dunning93}. Calculating a likelihood ratio in the general case involves considering all possible ways an $n$-gram could have been generated statistically from its component words: that is, whereas minLPR only considers the addition of one word to each $n-1$-gram, likelihood ratios must consider all combinations of all component $n-1,2,3\ldots$-grams, including, in our case, not only contiguous but also non-contiguous components. The corresponding exponential increase in complexity renders it and other such methods impractical for unrestricted FS identification.} we do not repeat all these comparisons here, but we do consider a lexicon built from ranking $n$-grams according to the measure used in our lattice (minLPR) as well as PMI. For each of these two association measures we build a lexicon equal to the size of the lexicon produced by our model. All comparisons here are subject to the same $n$-gram frequency threshold as the main model, since the content of our test sets are also linked to this threshold.

@@

The main results for FS acquisition across all four corpora are shown in \tabref{tab:main}. As noted in previous work, simple statistical association measures like PMI do fairly poorly when faced with unrestricted $n$-grams of variable length: minLPR is clearly a much better statistic for this purpose. The LPRseg method of Brooke et \al \shortcite{Brooke15b} consistently outperforms simple ranking, and the lattice method proposed here does better still, with a margin that is fairly consistent across languages. Generally, using clearing and overlap node interactions offer an relatively large increase in precision at the cost of a smaller drop in recall, though the change is fairly symmetrical in Croatian. The Japanese and ICWSM corpus had relatively high precision and low recall, whereas both the BNC and Croatian corpus have low precision and high recall.

@@

Given our interest in methods which work well on large corpora, the speed of the model is worth mentioning. Although considered as an individual process, optimizing the lattice is several orders of magnitude more time consuming than the decomposition heuristics of Brooke et \al \shortcite{Brooke15a}, we found that the time needed to build and optimize the lattice was still only a fraction of the time required to collect all the statistics necessary for to calculate LPR, a necessary first step for both methods: therefore, considered as end-to-end process, the effect of adding the lattice optimization is fairly minimal. We note that for all corpora, the lattice optimization algorithm converged within 10 iterations, and only the first two iterations require looking at majority of the nodes, since there is no need to revisit a node if nothing has changed in its vicinity.

>>>>

\begin{document}
\maketitle

\begin{abstract}
We present a new model for acquiring unrestricted multiword lexicons from large corpora based on  competition among $n$-gram candidates. In contrast to the standard approach of simple ranking by association measure, in our model $n$-grams are arranged in a lattice structure based on subsumption and overlap relationships, with nodes inhibiting other nodes in their vicinity when they are selected as a lexical item. We show how the configuration of such a lattice can be optimized tractably, and demonstrate using annotations of sampled $n$-grams that our method consistently outperforms alternatives by at least 0.05 F-score across several corpora and languages.
\end{abstract}

@@

Despite over 25 years of research in computational linguistics aimed at acquiring multiword lexicons using corpora statistics, and growing evidence that speakers both produce and process language primarily in terms of memorized sequences \cite{Wray08}, the individual word nonetheless stubbornly remains the \textit{de facto} standard processing unit for most research in modern NLP. The potential of multiword knowledge to improve both the automatic processing of language as well as offer new understanding of human acquisition and usage of language is the primary motivator of this work. We present a effective, expandable, and above all tractable new approach to comprehensive multiword lexicon acquisition that aims to find a middle ground between standard MWE acquisition approaches based on association measures \cite{Ramisch14}, and more sophisticated statistical models \cite{Newman12} which fail to scale to the large corpora which are the main sources of the distributional information in modern NLP systems.

@@

In this paper we will refer to the targets of our lexicon creation efforts as \termdef{formulaic sequences}, following the terminology of \newcite{Wray02,Wray08}, wherein a formulaic sequence (FS) is defined as ``a sequence, continuous or discontinuous, of words or other elements, which is, or appears to be, prefabricated: that is, stored and retrieved whole from memory at the time of use, rather than being subject to generation or analysis by the language grammar.'' In other words, a formulaic sequence shows signs of being part of a mental lexicon. Though by this definition single individuals or small groups may have their own FS, in this paper, we are only interested in FSs that are shared by a recognizable language community.

In computational linguistics, the most common term used to describe multiword lexical units is \emph{multiword expression} or ``MWE'' \cite{Sag02,Baldwin10}, but here we wish to make a principled distinction between at least somewhat non-compositional, strongly lexicalized MWEs and FS, a near superset which includes many MWEs but also (apparently) compositional linguistic formulas. This distinction is not a new one; it exists, for example, in the original paper of \newcite{Sag02} in the distinction between lexicalized and institutionalized phrases, and also to some extent in the MWE annotation of \newcite{Schneider14a}, who distinguish between weak (collocational)\footnote{Here we avoid the term \emph{collocation} entirely due to confusion with respect to its interpretation. Though in some contexts it has been used in a way that is similar to our definition of FS, it can be applied to any words that show a statistical tendency to appear in the vicinity of one another for any reason: for instance, the pair of words \ex{doctor/nurse} might be considered a collocation \cite{Ramisch14}.}  and strong (non-compositional) MWEs. It is our contention, however, that separate, precise terminology is useful for research targeted at either class: we need not strain the concept of MWE to include items which do not require special semantics, nor are we inclined to disregard the larger formulaticity of language simply because it is not the dominant focus of MWE research. Many MWE researchers would defensibly balk at including in their MWE lexicons and corpus annotations (English) formulaic sequences such as \ex{there is something going on}, \ex{it is more important than ever to \dotts}, \ex{\dotts do not know what it is like to \dotts}, \ex{there is no shortage of\dotts}, \ex{the rise and fall of\dotts}, \ex{now is not the time to\dotts}, \etc as well as tens of thousands of other such phrases which, along with less compositional MWEs like \ex{be worth \gap weight in gold}, fall under the FS umbrella. Another reason to introduce a different terminology is that there are classes of what is typically considered MWE that don't fit well into an FS framework, for instance novel compound nouns whose semantics are accessible by analogy (\eg \ex{plastic house}). Also, we exclude from the definition of both FS and MWE those named entities which refer to people or places which are little-known and/or whose surface form appears derived (\eg \ex{Mrs.\ Barbara W.\ Smith} or \ex{Smith Garden Supplies Ltd}). \figref{fig:terminology} shows the conception of the relationship between FS, MWE and (multiword) named entities that we assume for this paper.

@@

Regardless of the terminology used to describe them, the starting point for multiword lexicon creation has been typically lexical association measures \cite{Church90,Dunning93,Schone01,Evert04,Pecina10,DeAraujo11,Kulkarni11,Ramisch14}; when these methods are used to build a lexicon, particular syntactic patterns and thresholds for the metrics are typically chosen. Only some of these measures generalize tractably beyond two words, for example PMI \cite{Church90}, i.e.\ the log ratio of the joint probability to the product of the marginal probabilities of the individual words. Other measures specifically designed to address sequences of larger than two words include the $c$-value \cite{Frantzi00}, a metric designed for term extraction which weights term frequency by the log length of the $n$-gram while penalizing $n$-grams that appear in frequent larger ones, and mutual expectation \cite{Dias99}, which produces a normalized statistic that reflects how much a candidate phrase resists the omission of any particular word. Similarly, the lexical predictability ratio (LPR) of \newcite{Brooke15b} is an association measure intended for any possible syntactic pattern which is calculated by discounting syntactic predictability (as provided by part-of-speech information) from the overall conditional probability for each word given the other words in the phrase. Though most association measures involve only usage statistics of the phrase and its subparts, the DRUID measure is an exception which uses distributional semantics around the phrase to identify how easily an $n$-gram could be replaced by a single word \cite{Riedl15}.

Typically multiword lexicons are created by ranking $n$-grams according to an association measure and applying a threshold. The algorithm of \newcite{Silva99} is somewhat more sophisticated, in that it identifies the local maxima of association measures across subsuming $n$-grams within a sentence to identify MWEs of unrestricted length; its effectiveness beyond noun phrases, however, seems relatively limited \cite{Ramisch12}. \newcite{Brooke14a,Brooke15b} developed a heuristic method intended for general FS extraction in larger corpora, first using conditional probably statistics to do an initial (single pass) coarse-grained segmentation of the corpus, followed by a pass through the resulting vocabulary, breaking larger units into smaller ones based on a tradeoff between marginal and conditional statistics (\ie LPR). Beyond association measures, other general unsupervised approaches to the multiword unit identification include that of \newcite{Newman12}, who used a generative Dirichlet Process model which jointly creates a linear segmentation of the corpus and a multiword vocabulary of keyphrases. \newcite{Gimpel11} focus specifically on deriving word sequences with gaps using a generative model, with the intent of improving machine translation. A general drawback to these generative methods, relative to approaching involving association metrics, is scalability and lack of determinacy.

Other research in MWEs has tended to be rather focused on particular syntactic patterns, looking at, for instance, just verb/noun combinations \cite{Fazly09}. The work of \newcite{Schneider14b} is a rare example of a comprehensive token-level MWE identification system which distinguishes a full range of MWE sequences in the English Web Treebank, including those involving gaps, using a supervised sequence tagging model; like other models in this space, Schneider et \al make use of existing manual lexical resources and they note that an (unsupervised) automatic lexical resource could be useful addition to the model, though attempts to do so have had mixed results \cite{Riedl16}.

The motivation for building lexicons of formulaic sequences naturally overlaps with those for MWE: models of distributional semantics, in particular, can benefit from sensitivity to multiword units \cite{Cohen:Widdows:2009}, as can parsing \cite{Constant16} and topic models \cite{Lau+:2012b}. One major motivation for looking beyond MWEs is the ability to carry out broader linguistic analyses. Within corpus linguistics, multiword sequences have been studied in the form of \textit{lexical bundles} \cite{Biber04}, which are simply $n$-grams that occur above a certain frequency threshold: in order to get very reliable phrases, the threshold is typically set rather high (Biber et \al use 40 occurrences in 1 million words). Like FS, Lexical bundles generally involve larger phrasal chunks that would be missed by traditional MWE extraction, and so research in this area has tended to focus on how particular formulaic phrases (\eg \textit{if you look at}) are indicative of particular genres (\eg university lectures). Lexical bundles have been applied, in particular, to learner language: for example,  \newcite{Chen10} show that non-native student writers use a severely restricted range of lexical bundle types, and tend to overuse those types, while \newcite{Granger14} investigate the role of proficiency, demonstrating that intermediate learners underuse lower-frequency bigrams and overuse high-frequency bigrams relative to advanced learners. \newcite{Sakaguchi16} demonstrate that improving fluency (which is closely linked to the use of linguistic formulas) is more important than improving strict grammaticality with respect to native speaker judgments of non-native productions; \newcite{Brooke15b} explcitly argue for FS lexicons as a way to identify, track, and improve learner proficiency.

@@

The first step in the process is derive a set of $n$-grams and related statistics from a large, unlabeled corpus of text. Since our primary association measure is an adaption of LPR, our approach in this subsection mostly follows \newcite{Brooke15b} up until the last stage. An initial requirement of any such method is a $n$-gram frequency threshold, which for us (like Brooke et \al 2015) is 1 instance per 10 million words; we have done a manual analysis using the MWE corpus of \newcite{Schneider14a} which indicates very good (over 90\%) type-level MWE coverage using the frequency filtered $n$-gram statistics from the ICWSM blog corpus (see \secref{sec:evaluation}), excepting those highly particular proper names which we consider outside the scope of FS identification. Note that our set of $n$-grams also include gapped or non-contiguous $n$-grams; the definition of gaps is language specific---we will discuss the gaps for individual languages further in \secref{sec:evaluation}---but the general idea (and this is confirmed by empirical analysis) is that for any particular language there is a relatively small set of syntactic types of limited length that will tend to appear as a gap inside a formulaic sequence. In a part-of-speech tagged corpus, potential gaps can be identified by a part-of-speech regular expression. Putting in such a restriction is useful because it greatly lowers the number of gapped $n$-gram types, increasing precision and improving speed. Another way the number of types are lowered is by excluding punctuation and by lemmatizing.  Note that as long as the count threshold is significantly above 1, efficient extraction of all $n$-grams can be done iteratively ($n$ passes through the corpus) for both types in combination.

@@

In the segmentation approach of \newcite{Brooke15b}, LPR for an entire span is calculated as a product of the individual LPRs, but here we will use the minimum LPR across the words in the sequence (minLPR):

@@

\begin{displaymath}
E_0(t) = minLPR(t)^{-d_0(t)}
\end{displaymath}
And $E_1$ is the cost parameter $C$ taken to the negative power of an analogous function $d_1$:

\begin{displaymath}
E_1(t) = C^{-d_1(t)}
\end{displaymath}
The values of $d_0$ and $d_1$ are dependent on the context of the node in the lattice, but are both 1 when a node is not being influenced by other nodes. They are also restricted to always being positive.

@@

\begin{displaymath}
clear(t) = \prod_{i=1}^{m}{ min(1,\frac{minLPR(t)}{minLPR(u_i)})}
\end{displaymath}
where $u_i$ is the $i$th clearing node of $t$. We refer to this mechanism as clearing because it tends to clear away a variety of trivial uses of common FS which many have higher LPR due to the lexical and syntactic specificity of the FS. For instance, in \figref{fig:example} if the node \ex{keep \gap under wraps}  is turned on and has a minLPR of 8, then, if the minLPR of a node such as \ex{keep \gap under wraps for} is 4, clear(t) will be 0.5. We can now define $d_{0}(t)$ as:

@@

We evaluate our approach across three different languages including evaluation sets derived from four different corpora. In English, we follow \newcite{Brooke15b} in using a 890M token filtered portion of the ICWSM social media corpus \cite{ICWSM} tagged with the Tree Tagger \cite{Schmid95}. To facilitate a comparison with \newcite{Newman12}, which does not scale up to a corpus as large as the ICWSM, we also build a lexicon using the 100M token British National Corpus \cite{BNC}, using the standard CLAWS-derived POS tags for the corpus. Lemmatization included removing all inflectional marking from both words and POS tags. For English, we took the same POS regex used in \newcite{Brooke15b}, which includes simple nouns and portions thereof, up to a maximum length of 4.

@@

\newcite{Brooke15b} introduced a method for evaluating formulaic sequence extraction without a reference lexicon or direct annotation of the output of a model: instead, $n$-grams are sampled after applying the frequency threshold and annotated as being either an FS or not, allowing for calculation of a true F-score for any model. We use the annotation of 2000 $n$-grams in the ICWSM corpus from that earlier work, and applied the same annotation methodology to the other corpora discussed above: based on written guidelines derived from the definitions of Wray \cite{Wray08} and at least one practice round,three native-speaker, educated annotators judged 500 contiguous $n$-grams and another 500 gapped $n$-grams for each of the new corpora.

Other than the inclusion of new languages, our test sets differ from \newcite{Brooke15b} in two ways. First, instead of relying on strict majority annotation, we entirely excluded $n$-grams which just one annotator marked as FS. One advantage of a type-based annotation approach (as compared to a corpus annotation), particularly with regards to an task with a known subjective component, is that it is quite sensible to simply discard borderline cases, improving the reliability of evaluation at the cost of some representativeness.  Second, for the main evaluation we collapsed gapped and contiguous $n$-grams into a single test set. The rationale is that the number of positive gapped examples was too low (particularly for the BNC and the Croatian corpus) to provide a reliable independent F-score (see further of discussion of gapped phrases in \secref{sec:discussion}). Statistics for the four test sets are given in \tabref{tab:stats}.

@@

Our primary comparison is with the heuristic LPR model from \newcite{Brooke15b}, which is scalable to large corpora and includes gapped $n$-grams. For the BNC, we also compare, separately, the DP-seg model from \newcite{Newman12} with recommended settings, and the LocalMaxs algorithm of \newcite{Silva99} using the SCP as the association measure; because these other approaches only generate sequential multiword units, we use only the sequential part of the BNC test set. All comparison approaches have themselves been compared against a wide range of association measures in earlier work;\footnote{One popular association measure which is theoretically extensible to $n$-grams of any length but which has not been addressed in the context of unrestricted MWE/FS identification is the (log-)likelihood ratio \cite{Dunning93}. Calculating a likelihood ratio in the general case involves considering all possible ways an $n$-gram could have been generated statistically from its component words: that is, whereas minLPR only considers the addition of one word to each $n-1$-gram, likelihood ratios must consider all combinations of all component $n-1,2,3\ldots$-grams, including, in our case, not only contiguous but also non-contiguous components. The corresponding exponential increase in complexity renders it and other such methods impractical for unrestricted FS identification.} we do not repeat all these comparisons here, but we do consider a lexicon built from ranking $n$-grams according to the measure used in our lattice (minLPR) as well as PMI. For each of these two association measures we build a lexicon equal to the size of the lexicon produced by our model. All comparisons here are subject to the same $n$-gram frequency threshold as the main model, since the content of our test sets are also linked to this threshold.

@@

The main results for FS acquisition across all four corpora are shown in \tabref{tab:main}. As noted in previous work, simple statistical association measures like PMI do fairly poorly when faced with unrestricted $n$-grams of variable length: minLPR is clearly a much better statistic for this purpose. The LPRseg method of \newcite{Brooke15b} consistently outperforms simple ranking, and the lattice method proposed here does better still, with a margin that is fairly consistent across languages. Generally, using clearing and overlap node interactions offer an relatively large increase in precision at the cost of a smaller drop in recall, though the change is fairly symmetrical in Croatian. The Japanese and ICWSM corpus had relatively high precision and low recall, whereas both the BNC and Croatian corpus have low precision and high recall.

@@

Given our interest in methods which work well on large corpora, the speed of the model is worth mentioning. Although considered as an individual process, optimizing the lattice is several orders of magnitude more time consuming than the decomposition heuristics of \newcite{Brooke15a}, we found that the time needed to build and optimize the lattice was still only a fraction of the time required to collect all the statistics necessary for to calculate LPR, a necessary first step for both methods: therefore, considered as end-to-end process, the effect of adding the lattice optimization is fairly minimal. We note that for all corpora, the lattice optimization algorithm converged within 10 iterations, and only the first two iterations require looking at majority of the nodes, since there is no need to revisit a node if nothing has changed in its vicinity.
